A quantitative framework for investigating the reliability of empirical network construction

### 2019-01-22 manuscript/CI_interactions.tex
added Tomas changes

          Lastly, measurement errors are a pervasive source of uncertainty about ecological processes. Even a feasible interaction which occurs during sampling ($L_{ij}=1$, $X_{ij}=1$) may still go undetected~\citep{Jordano2016,Weinstein2017}. We can define the probability of detecting of an interaction, $D_{ij}$, given the interaction is feasible and occurs, as a stochastic process with  probability $P(D_{ij}|X_{ij}=1,L_{ij}=1)=\delta_{ij}$. Note that $D_{ij}$ is a stochastic variable where $D_{ij}=1$ if the interaction observed and $D_{ij}=0$ otherwise, while $P(D_{ij})$ is a continuous probability. Detection failure could happen for many reasons~\citep{Wirta2014}. Some sources of detection failure can be minimised (e.g., by combining multiple sampling methods~\citet{Wirta2014,Jordano2016}), but other sources are difficult to reduce. For example, it may not be possible to sample some species directly due to conservation concerns~\citep{Lagrue2015}. The probability of detecting an interaction also depends on the duration of the interaction and its evidence, the species involved, and the individuals sampled~\citep{Wells2013,Lagrue2015,Cirtwill2016,Weinstein2017}. As sampling improves $D_{ij}$ should tend towards 1 but, importantly, $D_{ij}$ will vary between interactions it is very likely that some $D_{ij}<1$. For example, when analysing gut contents interactions involving prey with hard parts will be more detectable than interactions with quickly-digested, soft prey.

@@

    In summary, an unobserved could be unfeasible [$P(L_{ij})=0$], be feasible but not occur during sampling ($P(X_{ij}|L_{ij}=1)=0$], or be feasible and occur during sampling but not be observed [$P(D_{ij}|X_{ij}=1,L_{ij}=1)=0$]. Researchers might be interested in identifying feasible interactions (i.e., reducing uncertainty about $\lambda_{ij}$) or identifying feasible interactions that occurred locally (i.e., reducing uncertainty about $\chi_{ij}$). Here we consider the former problem while noting that uncertainty about both $\chi_{ij}$ and $\delta_{ij}$ contribute to uncertainty about $\lambda_{ij}$. Future sampling should aim to characterise all three probabilities, as is possible in some studies focused on local realisations of a metaweb (e.g.,~\citet{Graham2018,Gravel2018}), but for now it is often difficult to distinguish between them.
    %One reason for this is the effect of factors such as abundance  on multiple levels of uncertainty. Rare species are less likely to interact at a given site due to neutral processes~\citep{Jordano2016,Graham2018} and interactions involving rare species are less likely to be detected~\citep{Bartomeus2013,Jordano2016}. Thus, abundance affects both $P(X_{ij})$ and $P(D_{ij})$.

@@

\section*{Materials and methods}

@@

        Note that, in this case, the prior distribution will be the same for all interaction probabilities. It is also possible to interpret other network properties (e.g., species degrees) in a similar network(s) as interaction probabilities and use these to create a prior (worked example in \emph{Appendix S4}). When using network summary statistics or degree distributions, it is important to consider whether the available information reflects what we believe about the network. Given the likelihood that most published networks are undersampled~\citep{Jordano2016}, it may be wise to include only the highest-quality networks available in a prior. Likewise, only the degree distribution for a similar system is likely to be a reasonable prior for a focal system. For example, if the focal system is large and contains a diverse array of species, it would be inappropriate to use a prior distribution drawn from a small network describing closely-related taxa (demonstrated in \emph{Appendix S7}). Similarly, highly-aggregated networks containing nodes which represent whole classes or families are likely to have very different structures than networks where most nodes represent single species as aggregated nodes are likely to have more interaction partners than single species.

@@

  \subsection*{An empirical example}

@@

\section*{Results}

@@

  \begin{figure}[h!]
      \caption{Despite the replication in our empirical dataset, most galler-natural enemy pairs were never observed co-occurring and those that co-occurred rarely interacted. \textbf{A)} Histogram of the number of pairs of species observed co-occurring at least once. 9794 galler-enemy pairs were never observed co-occurring (not shown).
      \textbf{B)} Histogram of the number of observed interactions between co-occurring species. Species which co-occurred but never interacted are included. 
      \textbf{C)} For each species pair, the number of observed interactions plotted against the number of observed co-occurrences. The red, dashed line indicates a 1:1 relationship.}
      \label{histograms}
      \begin{center}
      \includegraphics*[height=.6\textheight]{figures/GP_histogram.eps}
      % \includegraphics*[height=.6\textheight]{Figure_2.eps}
      \end{center}
      \end{figure}

>>>>

          Lastly, measurement errors are a pervasive source of uncertainty about ecological processes. Even a feasible interaction which occurs during sampling ($L_{ij}=1$, $X_{ij}=1$) may still go undetected~\citep{Jordano2016,Weinstein2017}. We can define the probability of detecting of an interaction, $D_{ij}$, given the interaction is feasible and occurs, as a stochastic process with  probability $P(D_{ij}|X_{ij}=1,L_{ij}=1)=\delta_{ij}$. Note that $D_{ij}$ is a stochastic variable where $D_{ij}=1$ if the interaction observed and $D_{ij}=0$ otherwise, while $P(D_{ij})$ is a continuous probability. Detection failure could happen for many reasons~\citep{Wirta2014}. Some sources of detection failure can be minimised (e.g., by combining multiple sampling methods~\citep{Wirta2014,Jordano2016}), but other sources are difficult to reduce. For example, it may not be possible to sample some species directly due to conservation concerns~\citep{Lagrue2015}. The probability of detecting an interaction also depends on the duration of the interaction and its evidence, the species involved, and the individuals sampled~\citep{Wells2013,Lagrue2015,Cirtwill2016,Weinstein2017}. As sampling improves $D_{ij}$ should tend towards 1 but, importantly, $D_{ij}$ will vary between interactions and it is very likely that some $D_{ij}<1$. For example, when analysing gut contents interactions involving prey with hard parts will be more detectable than interactions with quickly-digested, soft prey~\citep{Alberdi2018}.

@@

    In summary, an unobserved interaction could be unfeasible [$P(L_{ij})=0$], be feasible but not occur during sampling ($P(X_{ij}|L_{ij}=1)=0$], or be feasible and occur during sampling but not be observed [$P(D_{ij}|X_{ij}=1,L_{ij}=1)=0$]. Researchers might be interested in identifying feasible interactions (i.e., reducing uncertainty about $\lambda_{ij}$) or identifying feasible interactions that occurred locally (i.e., reducing uncertainty about $\chi_{ij}$). Here we consider the former problem while noting that uncertainty about both $\chi_{ij}$ and $\delta_{ij}$ contribute to uncertainty about $\lambda_{ij}$. Future sampling should aim to characterise all three probabilities, as is possible in some studies focused on local realisations of a metaweb (e.g.,~\citet{Graham2018,Gravel2018}), but for now it is often difficult to distinguish between them.
    %One reason for this is the effect of factors such as abundance  on multiple levels of uncertainty. Rare species are less likely to interact at a given site due to neutral processes~\citep{Jordano2016,Graham2018} and interactions involving rare species are less likely to be detected~\citep{Bartomeus2013,Jordano2016}. Thus, abundance affects both $P(X_{ij})$ and $P(D_{ij})$.

@@

\section*{Approach}

@@

        Note that, in this case, the prior distribution will be the same for all interaction probabilities. It is also possible to interpret other network properties (e.g., species degrees) in a similar network(s) as interaction probabilities and use these to create a prior (worked example in \emph{Appendix S4}). When using network summary statistics or degree distributions, it is important to consider whether the available information reflects what we believe about the network. Given the likelihood that most published networks are undersampled~\citep{Jordano2016}, it may be wise to include only the highest-quality networks available in a prior. Likewise, only the degree distribution for a similar system is likely to be a reasonable prior for a focal system. For example, if the focal system is large and contains a diverse array of species, it would be inappropriate to use a prior distribution drawn from a small network describing closely-related taxa (demonstrated in \emph{Appendix S7}). Similarly, data from a network where nodes are resolved to different taxonomic levels than the focal network are likely to have quite different properties and would not be an appropriate prior.

@@

  \section*{An empirical example}

@@

\subsection*{Empirical results}

@@

  \begin{figure}[h!]
      \caption{Despite the high replication in our empirical dataset, most galler-natural enemy pairs were never observed co-occurring and those that co-occurred rarely interacted. \textbf{A)} Histogram of the number of pairs of species observed co-occurring at least once. 9794 galler-enemy pairs were never observed co-occurring (not shown).
      \textbf{B)} Histogram of the number of observed interactions between co-occurring species. Species which co-occurred but never interacted are included. 
      \textbf{C)} For each species pair, the number of observed interactions plotted against the number of observed co-occurrences. The red, dashed line indicates a 1:1 relationship.}
      \label{histograms}
      \begin{center}
      \includegraphics*[height=.6\textheight]{figures/GP_histogram.eps}
      % \includegraphics*[height=.6\textheight]{Figure_2.eps}
      \end{center}
      \end{figure}
